Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-11-18 talk/pepm2011/paper.tex
I will ignore this comment
get space back. bit non-beautiful, but well.

\lstset{
  basicstyle=\setstretch{1.1}\ttfamily\footnotesize,
  language=Python,
  keywordstyle=\bfseries,
  stringstyle=\color{blue},
  commentstyle=\color{gray}\textit,
  fancyvrb=true,
  showstringspaces=false,
  %keywords={def,while,if,elif,return,class,get,set,new,guard_class}
  numberstyle = \tiny,
  numbersep = -20pt,
}

@@

\begin{figure}
\begin{lstlisting}[mathescape,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
class Base(object):
   pass

@@

\begin{figure}
\begin{lstlisting}[mathescape,numbers = right,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
# arguments to the trace: $p_{0}$, $p_{1}$
# inside f: res.add(y)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = int_add($i_{2}$, $i_{3}$)
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=10,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
$p_{5}$ = new(BoxedInteger)   |\setcounter{lstnumber}{11}|
set($p_{5}$, intval, $i_{4}$) |\setcounter{lstnumber}{14}|
$p_{6}$ = new(BoxedInteger)   |\setcounter{lstnumber}{16}|
set($p_{6}$, intval, -100)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=20,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
guard_class($p_{5}$, BoxedInteger)  |\setcounter{lstnumber}{21}|
$i_{7}$ = get($p_{5}$, intval)
guard_class($p_{6}$, BoxedInteger)  |\setcounter{lstnumber}{24}|
$i_{8}$ = get($p_{6}$, intval)
$i_{9}$ = int_add($i_{7}$, $i_{8}$)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=26,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
$i_{9}$ = int_add($i_{4}$, -100)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=44,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
$p_{15}$ = new(BoxedInteger)               |\setcounter{lstnumber}{45}|
set($p_{15}$, intval, $i_{14}$)            |\setcounter{lstnumber}{26}|
$p_{10}$ = new(BoxedInteger)               |\setcounter{lstnumber}{28}|
set($p_{10}$, intval, $i_9$)               |\setcounter{lstnumber}{54}|
jump($p_{15}$, $p_{10}$)
\end{lstlisting}

@@

\begin{figure}
\begin{lstlisting}[mathescape,numbers=right,escapechar=|, numberblanklines=false,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
# arguments to the trace: $p_{0}$, $p_{1}$ |\setcounter{lstnumber}{2}|
guard_class($p_1$, BoxedInteger)           |\setcounter{lstnumber}{4}|
$i_2$ = get($p_1$, intval)
guard_class($p_0$, BoxedInteger)           |\setcounter{lstnumber}{7}|
$i_3$ = get($p_0$, intval)
$i_4$ = int_add($i_2$, $i_3$)              |\setcounter{lstnumber}{25}|
$i_9$ = int_add($i_4$, -100)               |\setcounter{lstnumber}{35}|

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
$v^*$ = new($T_1$)
$w^*$ = new($T_2$)
set($w^*$, $L$, $u^*$)
set($w^*$, $R$, $u^*$)
set($v^*$, $L$, $w^*$)
set($v^*$, $R$, $v^*$)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,basicstyle=\setstretch{1.1}\ttfamily\scriptsize]
$w^*$ = new($T_2$)
set($w^*$, $L$, $u^*$)
set($w^*$, $R$, $u^*$)
\end{lstlisting}

@@

There exists a large number of works on escape analysis, which is a program
analysis that tries to find an upper bound for the lifetime of objects allocated
at specific program points
\cite{goldberg_higher_1990,park_escape_1992,choi_escape_1999,bruno_blanchet_escape_2003}.
This information can then be used to decide that certain objects can be
allocated on the stack, because their lifetime does not exceed that of the stack
frame it is allocated in. The difference to our work is that escape analysis is
split into an analysis and an optimization phase. The analysis can be a lot more
complex than our simple one-pass optimization. Also, stack-allocation reduces
garbage-collection pressure but does not optimize away the actual accesses to
the stack-allocated object. In our case, an object is not needed at all any
more. \reva{Has this relation pointed out before in some paper? Is it a
novel contribution of your work?}

>>>>

\lstset{
  basicstyle=\setstretch{1.05}\ttfamily\footnotesize,
  language=Python,
  keywordstyle=\bfseries,
  stringstyle=\color{blue},
  commentstyle=\color{gray}\textit,
  fancyvrb=true,
  showstringspaces=false,
  %keywords={def,while,if,elif,return,class,get,set,new,guard_class}
  numberstyle = \tiny,
  numbersep = -20pt,
}

@@

\begin{figure}
\begin{lstlisting}[mathescape,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
class Base(object):
   pass

@@

\begin{figure}
\begin{lstlisting}[mathescape,numbers = right,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
# arguments to the trace: $p_{0}$, $p_{1}$
# inside f: res.add(y)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = int_add($i_{2}$, $i_{3}$)
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=10,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$p_{5}$ = new(BoxedInteger)   |\setcounter{lstnumber}{11}|
set($p_{5}$, intval, $i_{4}$) |\setcounter{lstnumber}{14}|
$p_{6}$ = new(BoxedInteger)   |\setcounter{lstnumber}{16}|
set($p_{6}$, intval, -100)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=20,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
guard_class($p_{5}$, BoxedInteger)  |\setcounter{lstnumber}{21}|
$i_{7}$ = get($p_{5}$, intval)
guard_class($p_{6}$, BoxedInteger)  |\setcounter{lstnumber}{24}|
$i_{8}$ = get($p_{6}$, intval)
$i_{9}$ = int_add($i_{7}$, $i_{8}$)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=26,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$i_{9}$ = int_add($i_{4}$, -100)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,numbers=right,escapechar=|, firstnumber=44,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$p_{15}$ = new(BoxedInteger)               |\setcounter{lstnumber}{45}|
set($p_{15}$, intval, $i_{14}$)            |\setcounter{lstnumber}{26}|
$p_{10}$ = new(BoxedInteger)               |\setcounter{lstnumber}{28}|
set($p_{10}$, intval, $i_9$)               |\setcounter{lstnumber}{54}|
jump($p_{15}$, $p_{10}$)
\end{lstlisting}

@@

\begin{figure}
\begin{lstlisting}[mathescape,numbers=right,escapechar=|, numberblanklines=false,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
# arguments to the trace: $p_{0}$, $p_{1}$ |\setcounter{lstnumber}{2}|
guard_class($p_1$, BoxedInteger)           |\setcounter{lstnumber}{4}|
$i_2$ = get($p_1$, intval)
guard_class($p_0$, BoxedInteger)           |\setcounter{lstnumber}{7}|
$i_3$ = get($p_0$, intval)
$i_4$ = int_add($i_2$, $i_3$)              |\setcounter{lstnumber}{25}|
$i_9$ = int_add($i_4$, -100)               |\setcounter{lstnumber}{35}|

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$v^*$ = new($T_1$)
$w^*$ = new($T_2$)
set($w^*$, $L$, $u^*$)
set($w^*$, $R$, $u^*$)
set($v^*$, $L$, $w^*$)
set($v^*$, $R$, $v^*$)
\end{lstlisting}

@@

\begin{lstlisting}[mathescape,xleftmargin=20pt,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$w^*$ = new($T_2$)
set($w^*$, $L$, $u^*$)
set($w^*$, $R$, $u^*$)
\end{lstlisting}

@@

There exists a large number of works on escape analysis, which is a program
analysis that tries to find an upper bound for the lifetime of objects allocated
at specific program points
\cite{goldberg_higher_1990,park_escape_1992,choi_escape_1999,bruno_blanchet_escape_2003}.
This information can then be used to decide that certain objects can be
allocated on the stack, because their lifetime does not exceed that of the stack
frame it is allocated in. The difference to our work is that escape analysis is
split into an analysis and an optimization phase. The analysis can be a lot more
complex than our simple one-pass optimization. Also, stack-allocation reduces
garbage-collection pressure but does not optimize away the actual accesses to
the stack-allocated object. In our case, an object is not needed at all any
more.
